Analysis of Odd one out experiment

### 2020-04-24 reports/odd_one_out/analysis_oddoneout.Rmd
large changes in manuscript for review

```{r load libraries, warning=FALSE,message=FALSE}
set.seed(167)
library(tidyverse); theme_set(theme_classic(24))
library(lme4)
library(MuMIn)
library(here)

@@

@@

```{r visualize accuracy}
p <- df %>%
  group_by(subject_id, quintile) %>% 
  summarize(correct = mean(correct)) %>% 
  mutate(quintile = as.factor(quintile)) %>% 
  ggplot(aes(x = quintile, y = correct)) + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9) +
  ylab("Perc. correct") +

@@

p
ggsave(file.path(plots_dir, "Fig_chp3_results.svg"), p, width = 6, height = 6)

@@

p <- p0 + geom_point(data=predframe,colour="red")+
  geom_line(data=predframe,colour="red",aes(group=1)) + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9) +
  xlab("Distance ratio") + 
  ylab("Perc. correct") +
  theme(text = element_text(size = 24))

@@

```{r metrics results}
metrics <- metrics %>% filter(metric!= "ssim_correct")

p <- metrics %>% 
  filter(metric!= "ssim_correct" & metric!= "correct") %>% 
  ggplot(aes(x = quintile, y = value_fc7, col = metric, group = metric)) + 
  stat_summary(fun.data = "mean_cl_boot", position = position_jitter(0.2)) + 
  theme(aspect.ratio = 1) +
  coord_cartesian(ylim = c(0,0.5)) +
  geom_hline(yintercept = 1/9) +
  ylab("Perc. correct") +
  scale_x_continuous("Quintile", breaks = c(2,3,4), labels = c("2","3","4")) +
  theme(text = element_text(size=18)) + scale_color_manual(labels = c("GIST", "HOG", "RGB", "SIFT"), values = c("#E41A1C","#377EB8","#4DAF4A","#984EA3")) + 
  stat_summary(fun.y=mean, geom="line") +
  ggtitle("Predict CNN")
  #theme(legend.justification = c(1,1), legend.position=c(0.99,0.99))
p
ggsave(file.path(plots_dir, "Fig_chp3_metrics_explanation_fc7.svg"), p, width = 6, height = 6)

p <- metrics %>% 
  filter(metric!= "ssim_correct") %>% 
  ggplot(aes(x = quintile, y = value_partic, col = metric, group = metric)) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  coord_cartesian(ylim = c(0,0.5)) +
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9) +
  ylab("Perc. correct") +
  scale_x_continuous("Quintile", breaks = c(2,3,4), labels = c("2","3","4")) +
scale_color_manual(labels = c("Humans", "GIST", "HOG", "RGB", "SIFT"), values = c("#FF7F00","#E41A1C","#377EB8","#4DAF4A","#984EA3")) + 
  stat_summary(fun.y=mean, geom="line") +
  ggtitle("Predict human performance")
  #theme(legend.justification = c(1,1), legend.position=c(0.99,0.99))
p
ggsave(file.path(plots_dir, "Fig_chp3_metrics_predicthuman.svg"), p, width = 6, height = 6)

@@

@@

```{r test differences between metrics}
x <- metrics %>% filter(!is.infinite(value_partic)) %>% select(value_partic, subject_id,quintile,metric)

x[complete.cases(x),] %>% 
ez::ezANOVA(dv = value_partic, 
              wid = subject_id,
              within=.(quintile,metric))

x <- metrics %>% filter(!is.infinite(value_fc7)) %>% select(value_fc7, subject_id,quintile,metric,category)

x[complete.cases(x),] %>% 
ez::ezANOVA(dv = value_fc7, 
              wid = subject_id,
              within=.(quintile,metric))

@@

>>>>

```{r load libraries, warning=FALSE,message=FALSE}
set.seed(167)
library(tidyverse); theme_set(theme_classic(20))
library(lme4)
library(MuMIn)
library(here)
library(emmeans)

@@

metrics_fc7 <- readRDS(here("data", "oddoneout", "metrics_fc7_only_200407.rds"))

upperbound_subject <- readRDS(here("data", "oddoneout", "upperbound_subjects.rds"))
```

@@

```{r visualize accuracy}

p <- df %>%
  group_by(subject_id, quintile) %>% 
  summarize(correct = mean(correct)) %>% 
  mutate(quintile = as.factor(quintile)) %>% 
  ggplot(aes(x = quintile, y = correct)) + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9) +
  ylab("Perc. correct") +

@@

p
ggsave(file.path(plots_dir, "Figure1_results_left.svg"), p, width = 6, height = 6)

@@

p <- p0 + geom_point(data=predframe,colour="red")+
  geom_line(data=predframe,colour="red",aes(group=1)) + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9) +
  xlab("Distance ratio") + s
  ylab("Perc. correct") +
  theme(text = element_text(size = 24))

@@

```{r metrics results}
metrics <- metrics %>% filter(metric!= "ssim_correct")
metrics2 <- metrics %>% 
  select(metric, value_fc7, value_partic,value_fc7_top3,quintile) %>% 
  rbind(metrics_fc7 %>% mutate(metric = "baseline", value_fc7 = as.numeric(top1_outl_fc7 == target_position), value_fc7_top3 = top3_outl_fc7_correct, value_partic = NA) %>% select(metric, value_fc7, value_fc7_top3,value_partic,quintile))

metrics2 <- metrics2 %>% rbind(upperbound_subject %>% ungroup %>% mutate(value_partic = correct, value_fc7 = NA, value_fc7_top3 = NA, metric = "upper bound") %>% select(metric, value_fc7, value_partic,value_fc7_top3, quintile))

p <- 
  metrics2 %>% 
  filter(metric!= "correct", metric != "metrics_combined_correct", metric != "metrics_max_correct") %>% 
  ggplot(aes(x = quintile, y = value_fc7, col = metric, group = metric)) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  theme(aspect.ratio = 1) +
  coord_cartesian(ylim = c(0,1)) +
  geom_hline(yintercept = 1/9) +
  ylab("Perc. correct") +
  scale_x_discrete("Quintile", breaks = c(2,3,4), labels = c("2","3","4")) +
  theme(text = element_text(size=18)) + scale_color_manual(labels = c("upper bound", "GIST", "HOG", "RGB", "SIFT"), values = c("black","#E41A1C","#377EB8","#4DAF4A","#984EA3")) + 
  stat_summary(fun.y=mean, geom="line") +
  ggtitle("Predict CNN")
  #theme(legend.justification = c(1,1), legend.position=c(0.99,0.99))
p
ggsave(file.path(plots_dir, "Fig_chp3_metrics_explanation_fc7.svg"), p, width = 6, height = 6)
ggsave(file.path(plots_dir, "Fig_chp3_metrics_explanation_fc7_noleg.svg"), p+theme(legend.position="none"), width = 6, height = 6)

p <- metrics2 %>% 
  filter(metric!= "correct", metric != "metrics_combined_correct", metric != "metrics_max_correct") %>% 
  ggplot(aes(x = quintile, y = value_partic, col = metric, group = metric)) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  coord_cartesian(ylim = c(0,0.5)) +
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9) +
  ylab("Perc. correct") +
  scale_x_discrete("Quintile", breaks = c(2,3,4), labels = c("2","3","4")) +
scale_color_manual(labels = c("GIST", "HOG", "RGB", "SIFT","upper bound"), values = c("#E41A1C","#377EB8","#4DAF4A","#984EA3","black")) + # humans - "#FF7F00",
  stat_summary(fun.y=mean, geom="line") +
  theme(text = element_text(size=18))+
  ggtitle("Predict human performance")
  #theme(legend.justification = c(1,1), legend.position=c(0.99,0.99))
p
ggsave(file.path(plots_dir, "Fig_chp3_metrics_predicthuman.svg"), p, width = 6, height = 6)
ggsave(file.path(plots_dir, "Fig_chp3_metrics_predicthuman_noleg.svg"), p+theme(legend.position="none"), width = 6, height = 6)

p <- metrics %>% 
  ggplot(aes(x = quintile, y = value_partic_top3, col = metric, group = metric)) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  coord_cartesian(ylim = c(0,0.5)) +
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/3) +
  ylab("Perc. correct") +
  scale_x_continuous("Quintile", breaks = c(2,3,4), labels = c("2","3","4")) +
scale_color_manual(labels = c("GIST", "HOG", "RGB", "SIFT"), values = c("#E41A1C","#377EB8","#4DAF4A","#984EA3")) + 
  stat_summary(fun.y=mean, geom="line") +
  theme(text = element_text(size=18))+
  ggtitle("Predict human performance")
  #theme(legend.justification = c(1,1), legend.position=c(0.99,0.99))
p
ggsave(file.path(plots_dir, "Fig_chp3_metrics_predicthuman_top3.svg"), p, width = 6, height = 6)
ggsave(file.path(plots_dir, "Fig_chp3_metrics_predicthuman_top3_noleg.svg"), p+theme(legend.position="none"), width = 6, height = 6)

p <- metrics2 %>% 
  ggplot(aes(x = quintile, y = value_fc7_top3, col = metric, group = metric)) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  coord_cartesian(ylim = c(0,1)) +
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/3) +
  ylab("Perc. correct") +
  scale_x_discrete("Quintile", breaks = c(2,3,4), labels = c("2","3","4")) +
theme(text = element_text(size=18)) + 
  scale_color_manual(labels = c("upper bound", "GIST", "HOG", "RGB", "SIFT"), values = c("black","#E41A1C","#377EB8","#4DAF4A","#984EA3")) + 
  stat_summary(fun.y=mean, geom="line") +
  ggtitle("Predict CNN")
  #theme(legend.justification = c(1,1), legend.position=c(0.99,0.99))
p
ggsave(file.path(plots_dir, "Fig_chp3_metrics_explanation_fc7_top3.svg"), p, width = 6, height = 6)
ggsave(file.path(plots_dir, "Fig_chp3_metrics_explanation_fc7_top3_noleg.svg"), p+theme(legend.position="none"), width = 6, height = 6)
```

@@

metrics %>%
  filter(metric == "metrics_combined_correct") %>% 
  group_by(quintile) %>% 
  filter(!is.infinite(value_fc7)) %>% 
  summarize(value_fc7 = mean(value_fc7, na.rm = T),
            value_partic = mean(value_partic, na.rm = T))
```

```{r test differences between metrics}

aov_value_partic <- 
  metrics %>% filter(metric!= "correct", metric != "metrics_combined_correct", metric != "metrics_max_correct") %>% 
  mutate(quintile = as.character(quintile)) %>% 
lme4::lmer(value_partic ~ quintile*metric + (1|subject_id),.)

emm_partic_top1 <- emmeans(aov_value_partic, ~ metric | quintile, lmer.df = "asymptotic")
eff_partic_top1 <- eff_size(emm_partic_top1, sigma = sigma(aov_value_partic), edf = df.residual(aov_value_partic))

aov_value_fc7 <- 
  metrics %>% filter(metric!= "correct", metric != "metrics_combined_correct", metric != "metrics_max_correct") %>% 
  mutate(quintile = as.character(quintile)) %>% 
lme4::lmer(value_fc7 ~ quintile*metric +(1|subject_id),.)

emm_fc7_top1 <- emmeans(aov_value_fc7, ~ metric | quintile, lmer.df = "asymptotic")
eff_fc7_top1 <- eff_size(emm_fc7_top1, sigma = sigma(aov_value_fc7), edf = df.residual(aov_value_fc7))

aov_value_partic_top3 <- 
  metrics %>% filter(metric!= "correct", metric != "metrics_combined_correct", metric != "metrics_max_correct") %>% 
  mutate(quintile = as.character(quintile)) %>% 
lme4::lmer(value_partic_top3 ~ quintile*metric + (1|subject_id),.)

emm_partic_top3 <- emmeans(aov_value_partic_top3, ~ metric | quintile, lmer.df = "asymptotic")
eff_partic_top3 <- eff_size(emm_partic_top3, sigma = sigma(aov_value_partic_top3), edf = df.residual(aov_value_partic_top3))

aov_value_fc7_top3 <- 
  metrics %>% filter(metric!= "correct", metric != "metrics_combined_correct", metric != "metrics_max_correct") %>% 
  mutate(quintile = as.character(quintile)) %>% 
lme4::lmer(value_fc7_top3 ~ quintile*metric +(1|subject_id),.)

emm_fc7_top3 <- emmeans(aov_value_fc7_top3, ~ metric | quintile, lmer.df = "asymptotic")
eff_fc7_top3 <- eff_size(emm_fc7_top3, sigma = sigma(aov_value_fc7_top3), edf = df.residual(aov_value_fc7_top3))

rbind(eff_partic_top1 %>% as.data.frame(),
      eff_fc7_top1 %>% as.data.frame()) %>% 
  pull(effect.size) %>% abs() %>% range()
rbind(eff_partic_top3 %>% as.data.frame(),
      eff_fc7_top3 %>% as.data.frame()) %>% 
  pull(effect.size) %>% abs() %>% range()

@@

### Test metrics on chance level

```{r test metrics on chance level}
 metrics %>% filter(metric!= "correct", metric != "metrics_combined_correct", metric != "metrics_max_correct") %>% 
  filter(quintile == 2) %>% 
  group_by(metric) %>% 
  do(t.test(x = .$value_fc7_top3, mu = 1/3) %>% broom::tidy())
```

@@

df %>% 
  group_by(quintile,subject_id) %>% 
  summarize(correct = mean(correct)) %>% 
  arrange(-correct)
```
